---
title: "Slopes-colored"
author: "Joachim Goedhart"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Colored slopes

The purpose of this document is to demonstrate how paired data can be colored based on the slope of the connecting line. In this example, a negative slope will be blue and a positive slope red.

Let's load the necessary library first:

```{r}
library(tidyverse)
```

We'll create some data that has the columns 'replicates', 'Condition' and a measured 'value':

```{r}
df <- data.frame(replicate = rep(1:50, 2), Breast_tissue = c(rep("Healthy",50),rep("Tumor",50)), fprot=runif(100))
head(df)
```

Next, we will calculate the difference for `fprot` between the  "Tumor" and "Healthy" values per replicate, as this defines the trend (positive or negative):

```{r}
df <- df %>% group_by(replicate) %>%
         mutate(difference = (fprot[Breast_tissue=="Tumor"] - fprot[Breast_tissue=="Healthy"]))
head(df)

```

First we generate the slope plot:

```{r}
ggplot(df, aes(x=Breast_tissue, y=fprot)) + geom_point() + geom_line(aes(group=replicate))
```

Now, with colored lines that are based on the value for "difference' (and some further styling of the layout):

```{r}
p <- ggplot(df, aes(x=Breast_tissue, y=fprot)) +
  geom_line(aes(group=replicate, color=difference>0), linewidth=1, alpha=0.6) +
  geom_point(color = "black", fill="grey", shape=21, size=3, alpha=0.4)

p <- p + theme_light(base_size = 16)
p
```

To define the colors manually:

```{r}
p <- p + scale_color_manual(values = c("darkblue","red"))
p
```

